---
title: "Quantifying effect of planting dates on potential potato yield"
subtitle: "A crop modeling approach"
author: "Johan Ninanya (noni)"
date: "`r Sys.Date()`"
#site: bookdown::bookdown_site
#documentclass: book
output:
  rmdformats::readthedown:
    highlight: kate
    number_sections: FALSE
    code_folding: show
    code_download: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Background

This notebook aims to document the simulation exercise performed by Ramírez et al. (2024). The [SOLANUM](https://cipotato.org/site/inrm/home/downmod.htm) model was used to determine the potential potato yield (Yp) and "roughly" quantify the effect of planting dates on Yp. The model was calibrated using the [SOLANUM's Parameter Estimator](https://doi.org/10.1515/opag-2018-0019) and fed with weather data from [NASA POWER](https://cran.r-project.org/web/packages/nasapower/index.html). Please see more details in Ramírez et al. (2024).

## Crop model calibration

For this, the **SOLANUM's Parameter Estimator** was used. A complete documentation about this tool can be found in [Harahagazwe et al. (2018)](https://doi.org/10.1515/opag-2018-0019)

```{r cars}
# load libraries
library(nasapower)
library(meteor)
library(lubridate)
library(dplyr)

load(url("https://github.com/jninanya/Ramirez-et-al-2024/raw/main/solanumR/CropParamsList.Rdata"))

CropParamsList$BariAlu72

# defining variables (wvars) and period
wvars <- c("T2M_MAX", "T2M_MIN", "ALLSKY_SFC_SW_DWN")
period <- c("2000-01-01", "2023-12-31")

# coordinates (lon, lat) of the Fltola location
FUL <- c(89.5262, 22.7088) 

# getting daily data from NASA POWER 
wdata1 <- get_power(
  community = "ag",
  lonlat = FUL,
  pars = wvars,
  dates = period,
  temporal_api = "daily"
)
```

# Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
